---
title: "above and belowground sampling -Stat summary-"
author: "Shun Hasegawa"
date: "10/27/2016"
output:
  html_document:
    toc: true
---

```{r setup, include=FALSE}
library(car)
library(lmerTest)
library(ggplot2)
library(grid)
library(pander)
library(plyr)
library(dplyr)
library(tidyr)
library(vegan)

knitr::opts_chunk$set(echo = TRUE)
load("../Output/Data/all_obj.RData")
theme_set(theme_bw()) 

# function to generate anova table from lmer
lmer_anova_tbl <- function(lmermod, ...){
  a <- Anova(lmermod, test.statistic = "F")
  panderOptions('round', c(2, 0, 0, 3))
  if(any(a[, 4] <= 0.05)) emphasize.strong.cells(cbind(which(a[, 4] <= 0.05), 4))
  emphasize.italics.cols(4)
  pander(a,  justify = "right", ...)
}


lm_anova_tbl <- function(lmmod, ...){
  a <- Anova(lmmod)
  ap <- a$`Pr(>F)`[!is.na(a$`Pr(>F)`)]
  panderOptions('round', c(2, 0, 2, 3))
  if(any(ap <= 0.05)) emphasize.strong.cells(cbind(which(a[, 4] <= 0.05), 4))
  emphasize.italics.cols(4)
  pander(a,  justify = "right", ...)
}


```

## Introduction
Here, I show the sumamry results of anlaysis on for "hole" sampling carried out in September 2015 from the DRIGRASS experiment. The hole will referred to as subplot, hereafter. For a community-level (or subplot-level), I anlysed it in the same way as the harvest data. Firstly, interaction effects between rainfall and herbivore treatments were assessed. Where there was no herbivore effect, the complete dataset, including added herbivore plots, was analysed. The following variables are analysed.

* total biomass
* above : belowground biomass ratios
* Belowground biomass at 0-10cm : 10-20cm ratios

For a species-level (i.e. plant traits), the herbivore treatment was not considered as there was not sufficient replicates. There were 20 root diameter classes. Cumulative sum at each diameter class was computed prior to analysis. The values at a diameter class of 20 were not used as they occasionally contain odd values. Firstly, rainfall treatments effects were analysed for each cumulative diameter class for each species. Secondly, differences in traits between species were assessed. The following plant traits were evaluated.

* root length by root mass
* root surface area by root mass
* number of root forks (or root branching) by root mass
* number of root tips (or root ends) by root mass
* stem:root ratios

## Community biomass

```{r comm_biom_fig, echo=FALSE, fig.height=4, fig.width=6, results='hide', warning=FALSE, fig.cap="Community biomass (mean±SE) for above (ab_mass), below at 0-10 cm (bl_mass_0_10) and below at 20-20 cm (bl_mass_10_20) biomass"}
fig_comm_biom
```

### Total community biomass
```{r comm_biom_ttl_biom_tbl, echo=FALSE, results='asis', message=FALSE, fig.show='hold'}

lm_anova_tbl(ttlb_rxh_m1, caption = "Total biomass for rainfall and herbivore treatments")
lm_anova_tbl(ttlb_r_m1, caption = "Total biomass for rainfall treatments")

```

### above:below-ground biomass ratio

```{r comm_biom_ab_ratio_tbl, echo=FALSE, results='asis', message=FALSE, fig.show='hold'}

lm_anova_tbl(abr_rxh_m1, caption = "Above:Belowground ratios for rainfall and herbivore treatments")
lm_anova_tbl(abr_r_m1, caption = "Above:Belowground ratios for rainfall treatments")

```

### Belowground biomass at 0-10cm : 10-20cm ratios

```{r comm_biom_bl_10_20_ratio_tbl, echo=FALSE, results='asis', message=FALSE, fig.show='hold'}

lm_anova_tbl(bl12r_rxh_m1, caption = "Belowground biomass at 0-10 : 10-20cm ratios for rainfall and herbivore treatments")
lm_anova_tbl(bl12r_r_m1, caption = "Belowground biomass at 0-10 : 10-20cm ratios for rainfall treatments")

```


## Plant traits

### Root length

#### Rainfall treatment

```{r plant_trait_rootL_byRain_fig1, echo=FALSE, fig.height=6.5, fig.width=6.5, results='hide', warning=FALSE, fig.cap="Root length at each diameter class"}
ggplot(prcssd_root_LSA2, aes(x = dmclass, y = L , col = treatment))+
  labs(y = expression(Root~length~(mm~mg^'-1'))) +
  facet_wrap(~ spp, scale = "free_y") +
  geom_point(size = 1, alpha = .4) +
  geom_smooth(se = FALSE) +
  theme(legend.position = "top")
```

```{r plant_trait_rootL_byRain_fig2, echo=FALSE, fig.height=6.5, fig.width=6.5, results='hide', warning=FALSE, fig.cap="Cumulative root length at each diameter class"}
ggplot(prcssd_root_LSA3, aes(x = dmclass, y = L.cum, col = treatment))+
  labs(y = expression(Cumulative~root~length~(mm~mg^'-1'))) +
  facet_wrap(~ spp, scale = "free_y") +
  geom_point(size = 1, alpha = .4) +
  geom_smooth(se = FALSE) +
  theme(legend.position = "top")

```

#### Species comparison

```{r plant_trait_rootL_bySpp_fig, echo=FALSE, fig.height=4, fig.width=6.5, results='hide', warning=FALSE, fig.cap="Total root length for each species"}
filter(prcssd_root_LSA3, dmclass == 19) %>%
  ggplot(., aes(x = spp, y = log(L.cum))) +
  labs(y = "log(Total root length) (mm mg-1)") +
  geom_boxplot() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))

```

### Root surface area

#### Rainfall treatment
```{r plant_trait_rootSA_byRain_fig1, echo=FALSE, fig.height=6.5, fig.width=6.5, results='hide', warning=FALSE, fig.cap="Root surface area at each diameter class"}
ggplot(prcssd_root_LSA4, aes(x = dmclass, y = SA , col = treatment))+
  labs(y = expression(Root~SA~(mm^2~mg^'-1'))) +
  facet_wrap(~ spp, scale = "free_y") +
  geom_point(size = 1, alpha = .4) +
  geom_smooth(se = FALSE) +
  theme(legend.position = "top")
```


```{r plant_trait_rootSA_byRain_fig2, echo=FALSE, fig.height=6.5, fig.width=6.5, results='hide', warning=FALSE, fig.cap="Cumulative root surface area at each diameter class"}
ggplot(prcssd_root_LSA5, aes(x = dmclass, y = SA.cum, col = treatment))+
  labs(y = expression(Root~SA~(mm^2~mg^'-1'))) +
  facet_wrap(~ spp, scale = "free_y") +
  geom_point(size = 1, alpha = .4) +
  geom_smooth(se = FALSE) +
  theme(legend.position = "top")

```

#### Species comparisons

```{r plant_trait_rootSA_bySpp_fig, echo=FALSE, fig.height=4, fig.width=6.5, results='hide', warning=FALSE, fig.cap="Total root surface area for each species"}
filter(prcssd_root_LSA5, dmclass == 19) %>%
  ggplot(., aes(x = spp, y = log(SA.cum))) +
  labs(y = "log(Total root surface area) (mm2 mg-1)") +
  geom_boxplot() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))

```




### Root forks

#### Rainfall treatment
```{r plant_trait_rootForks_byRain_fig, echo=FALSE, fig.height=6.5, fig.width=6.5, results='hide', warning=FALSE, fig.cap="Root forks"}

ggplot(raw_root_trait_byDC, aes(x = treatment, y = Forks, fill = treatment)) +
  labs(y = "mg-1") +
  facet_wrap(~ spp, scale = "free_y") +
  geom_boxplot() +
  theme(legend.position = "top")

```

#### Species comparison

```{r plant_trait_rootForks_bySpp_fig, echo=FALSE, fig.height=4, fig.width=6.5, results='hide', warning=FALSE, fig.cap=" Number of root forks for each species"}
filter(raw_root_trait_byDC) %>%
  ggplot(., aes(x = spp, y = log(Forks))) +
  labs(y = "log(Number of root forks) (mg-1)") +
  geom_boxplot() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))

```




### Root tips

#### Rainfall treatment

```{r plant_trait_rootTips_byRain_fig, echo=FALSE, fig.height=6.5, fig.width=6.5, results='hide', warning=FALSE, fig.cap="Root tips"}

ggplot(prcssd_root_trait_byDC1, aes(x = treatment, y = Tips, fill = treatment)) +
  labs(y = "mg-1") +
  facet_wrap(~ spp, scale = "free_y") +
  geom_boxplot() +
  theme(legend.position = "top")


```

#### Species comparison

```{r plant_trait_rootTips_bySpp_fig, echo=FALSE, fig.height=4, fig.width=6.5, results='hide', warning=FALSE, fig.cap=" Number of root forks for each species"}
filter(prcssd_root_trait_byDC1) %>%
  ggplot(., aes(x = spp, y = log(Tips))) +
  labs(y = "log(Number of root tips) (mg-1)") +
  geom_boxplot() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))

```


### Stem:Root ratios

#### Rainfall treatment

```{r plant_trait_srR_byRain_fig, echo=FALSE, fig.height=6.5, fig.width=6.5, results='hide', warning=FALSE, fig.cap="Stem:Root ratio", , warning=FALSE}

ggplot(prcssd_root_trait_byDC1, aes(x = treatment, y = sr_ratio, fill = treatment)) +
  facet_wrap(~ spp, scale = "free_y") +
  geom_boxplot() +
  theme(legend.position = "top")


```


#### Species comparison

```{r plant_trait_srR_bySpp_fig, echo=FALSE, fig.height=4, fig.width=6.5, results='hide', warning=FALSE, fig.cap=" Number of root forks for each species", warning=FALSE}
filter(prcssd_root_trait_byDC1) %>%
  ggplot(., aes(x = spp, y = log(sr_ratio))) +
  labs(y = "log(Stem:Root ratios)") +
  geom_boxplot() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))

```




### Sumamry stats for plant traits


#### Rainfall treatment
```{r plant_trait_byRrain_smmry_tbl, echo=FALSE, warning=FALSE}
a <- smmry_anova_byRain_tbl
emphasize.italics.cols(5)
emphasize.strong.cells(cbind(which(a[, 5] <= 0.05), 5))
pander(a,  justify = "right", split.table = Inf, caption = "Results of one-way ANOVA for rainfall effects and summary with mean±SE. SE is NA when n < 2. Where there was a significant treatment effect at P < 0.05, post-hoc tests were performed. The results of post-hoc comparisons are shown by letters alongside parenthesis.")

```


#### Species comparison
```{r palnt_trait_bySpp_smmry_tbl, echo=FALSE, warning=FALSE}
a <- smmry_anova_bySpp_tbl %>% 
  gather(spp, value, -trait, -unit, -F, -P) %>% 
  gather(variable, value2, -spp, -trait) %>% 
  spread(trait, value2) %>% 
  filter(!(spp != "Axonopus" & variable %in% c("F", "P", "unit"))) %>% 
  mutate(variable = ifelse(variable == "value", spp, variable)) %>% 
  select(-spp)

pander(a,  justify = "right", split.table = Inf, caption = "Results of one-way ANOVA for species comparisons and summary with mean±SE. SE is NA when n < 2. Where there was a significant treatment effect at P < 0.05, post-hoc tests were performed. The results of post-hoc comparisons are shown by letters alongside parenthesis.")


```

